### 2009-08-27 docs/pubs/ISEI6/response.txt
start of responses and edits

Mark's Revisions:

On p. 7, the "sufficient blocks" are not given a template
representation, whereas all the other blocks are...

Also on p. 7, you describe how the relationship block can be negated
in a sufficient block.  Is negation also allowed in the relationship
block itself?  This would be nice, and if permissible, should be
described in the relationship block section.

==> add to possible future extensions

Section 3.2 on reasoning is where some additional explanation could be
provided about difficulties of adding domain/range constraints to an
"advanced owlifier" .  Also, does the relationship block
simultaneously imply both someValuesFrom AND allValuesFrom (in
Manchester syntax)?  If not, it would be nice to provide a closure
axiom as an option on the relationship block.  Is this low-hanging
fruit?

==> add to possible future extensions

In second paragraph of conclusion, last sentence, I am not sure why
you refer to "disjoint classes" as a form of global taxonomic
constraint defined in [22]-- I would think it is more the converse
that global taxonomic constraints often require disjoint classes.

==> add to possible future extensions

It would be nice to clearly say what currently DOES or DOES NOT YET
exist in owlifier, either in the Conclusion, or the preceding
paragraph.

==> add to implementation part, along with a url to the OWLifier webpage

----------------------------------------------------------------------
Reviewer #1: 
----------------------------------------------------------------------
1). "Not mentioned is the list of OWL features not supported."

- Probably add a list to the implementation?

[josh]: 
Features not supported:
  - Data-type properties
  - Property domains and ranges
  - Symmetric properties
  - Others?
  I would say at page 2, third full paragraph, after first sentence:
  Primarily, we omit much of the complexities associated with
  properties (e.g., domains and ranges, data-type properties) that are
  often redundant and should be implemented where necessary, by more
  experienced knowledge modellers.

[mark]: I'd amend Josh's sentence slightly, and place it as the last
sentence of the preceding paragraph (para. 2; before paragraph
starting with "While not as expressive") rather than second sentence
of following para:

"We omit several OWL features associated with properties (e.g.,
domains and ranges, data-type properties) as these can be confusing to
construct, and more suitable for experienced knowledge modelers."

----------------------------------------------------------------------
2). "Overriding the OWL open world assumption should be fine, so long as
that assumption is made clear to the users int eh documentation."

We primarily provide a different approach for specifying disjoint
classes, as opposed to overriding the OWA in general. However, we
agree that describing this feature will be particularly important to
users of OWLifier that are familiar with OWL-DL.

[josh]:
- This is currently made clear in section 3.1 Disjoint Class Inference
- Maybe there should be a brief mention of this in the introduction?

[mark]: I think it is fine as is. Hard to be brief about this without
being confusing.

----------------------------------------------------------------------
3). "There is no mention of the competing CMAP-COE tool that makes it
easy to create ontologies visually, with help from drop-down lists.
It would be helpful to include a line that compares your approach to
CMAP (which also supports only a subset of OWL)."

We added additional text in the introduction that further describes
our motivation for the spreadsheet-based approach and some of its
advantages for ecologists/sciencts over CMAP-COE.

[josh]:

- Even installing and using this tool is too complicated and a pain
for people. Introduction, third paragraph, second sentence, maybe
change to something like:

Our approach is to allow scientists to use common and familiar
spreadsheet-based tools (e.g., Microsoft Excel, Apple Numbers, Open
Office Spreadsheet) to describe, in an intuitive way, different
aspects of an ontology, and then to take these descriptions and
convert them into full-fledged OWL ontologies using a software
application called owlifier.  The use of a familiar editing
environment will improve the speed and understanding of ontology
construction and avoid locating, downloading, and interpreting
additional software packages and add-ons (e.g., CMAP-COE).  Many
spread-sheet applications also contain easy-to-use features that
support scientists with the structuring process (e.g., cut / copy /
paste, search / replace, sort, split-window, annotation,
track-changes, freeze panes, and so on).

[mark]: Above sounds good to me.

----------------------------------------------------------------------
4). "Something that might be pointed out - the tool could be used to
help enforce some community standards (e.g., a common use of a
"part-of" relation)."

[josh]:
- It would be nice to say something like (although I not sure where it
  would be placed):

Furthermore, blocks can be reused (e.g., relationship blocks) to
ensure that future ontology development adheres to any community
agreed-upon representations, and will result in the smoother ontology
integration within ecology.

[mark]: Nice

[shawn]: Because owlifier spreadsheets can be converted to owl
ontologies, these can be further imported in owlifier spreadsheets
(e.g., for extension).  This follows the basic OWL infrastructure.

----------------------------------------------------------------------
5). "In the references: Usually, the word "pages" is abbreviated to
pp. Ref 18: should be to a journal article, not a web site - perhaps
Raskin and Pan (2005) in Computers & Geosciences. Ref 19: to EKAW?
What is that?  Compare Refs 4 and 7 - different wordings are used to
describe the same conference."

FIXED

----------------------------------------------------------------------
Reviewer #2:
----------------------------------------------------------------------
6). "In the early stages of the paper there are some assertions
without reference about what is and is not difficult. See p.2 end of
2nd, 3rd and 4th paragraphs.  Can a little more detail be added to
support the statements? If not, then noting that these are
suppositions would be useful for the reader."

[josh]: Honestly, I don't find a problem here.  I think paragraph 2 is
fine and obvious to anyone who's used Protege (we could potential say
"based on our experience in an ecological informatics project
constructing ontologies [SEEK]").  I can't find anything not
referenced in paragraph 3.  And the example used by the reviewer in
paragraph 4 IS referenced and not so much about the level of logical
description, and basically says "you can capture the level of
description shown in this cited paper as necessary for improved
discovery and integration".

[mark]: I think the assertion about "lack of suitable ontology editing
tools is a major barrier for widescale adoption in ecology" is the
issue, and begs the question as to why this isn't the case in other
domains.  So, I suggest we delete the last sentence of paragraph 2 on
pg 2, and add a following paragraph that elaborates and clarifies that
point:

"A well-funded ontology development effort, Gene Ontology (GO) has
been underway for over a decade within the molecular biology
community, and has led to some major successes enriching genomic and
biomedical data resources with annotations to a community-based
knowledge representation. A major lesson learned from these efforts
was the importance of keeping the ontology focused, and indeed GO has
a highly delimited scope (Bada et al.)  Ecology, on the other hand,
encompasses an extremely broad range of scales and disciplines, such
that a coordinated ontology development effort would need to involve a
large number of experts to cover the diverse types of information
relevant to ecological investigations.  At this point, we believe that
ecological researchers need to begin constructing knowledge
representations for their own specialized subdisciplines.  Most
ontology editing tools, however, typically require substantial
expertise in formal knowled ge representation, and present a major
barrier for more wide-scale creation and adoption of ontologies in
ecology."

[Bada, Michael, Robert Stevens, Carol Goble, A. Gil, Michael
Ashburner, Judith A. Blake, J. Michael Cherry, Midori Harris, Suzanna
Lewis.  2004. A short study on the success of the gene ontology.
Journal of Web Semantics, vol 1: 235-240.]

----------------------------------------------------------------------
7). "Was there any quantifiable basis for choosing the predicates?  In
particular, the statement" "While not as expressive as OWL-DL, our
approach can produce ontology structures that are essential for
improved data discovery and integration [14]."  indicates some level
of decision in the reduced description but without any details,
especially in terms for DL-expressivity, it is SHIQ, SHOIN, or what?"

As far as I am aware, the arrangement of DL constructs used in
OWLifier does not directly map to any previously named DL sublanguage,
however, it does provides only OWL-DL based constructs, and thus can
be considered best as a sublanguage of SHOIN(D). Note that although
many of the OWL-DL constructs can be used, they can only be used in
certain contexts. Also, some of the more foundational constructs of
DLs are not considered, e.g., value restrictions and concept
disjunction. While additional blocks can easily be added to OWLifier
spreadsheet conventions, the proposed set of blocks were chosen to
provide a simplified syntax (for a novice DL user) while leading to a
"natural" or "intuitive" semantics of blocks for these users.

----------------------------------------------------------------------
8). "There are no mention of instances, that I could find, at all of
instances?  Are these part of the use case or not? If not, this really
needs to be stated.  An ontology without instances has limited
application."

[josh]: Maybe at the start of third paragraph of introduction: This
paper presents a novel approach for ontology creation that aims at
being more intuitive for ecologists that can be used to rapidly
construct large ontologies of ecological terms (classes) and the
subsequent annotation of scientific data (instances).

----------------------------------------------------------------------
9). "Fig.1 suggests a round-trip approach to knowledge engineering,
but in fact it is not. How does the present formulation represent to
scientiest what gets dropped in the back conversion from OWL to the
spreadsheet format?"

[josh]: I guess we need be a bit more transparent in the last sentence
of the bottom paragraph, page 10; something like:
   While straightforward to perform, this last conversion will only
preserve OWL-DL constructs that are represented by owlifier blocks.

[mark]: Substitute term "back conversion" instead of "last
conversion", and "that are" to "that can be"

Note misspelling of "appraoches" on p. 10.

----------------------------------------------------------------------
10). "Most of all, there is almost no (or none at all) description of
the implemenation or any evaluation, i.e. the how, why, etc.  What has
this application been applied to?"

[josh]: We could mention that current efforts are underway using
owlifier to structure list of keywords from the US LTER program.  And
that the spread-sheet approach works well for these kind of mid-level
term sorting exercises (e.g., 50-5000) terms.  Nonetheless, (from
experience) larger numbers of terms does get difficult to navigate and
structure, although the use of blocks means that you can focus on
tractable chunks of terms and properties.  Furthermore, the use of
spread-sheet software features greatly improve navigation, commenting,
and editing capabilities (e.g., sort, freeze panes, split window,
search, multiple worksheets / windows, comments, track-changes).

----------------------------------------------------------------------
11). "It seems that it would be easy to generate complex blocks. How
are moderate size ontologies developed?"

----------------------------------------------------------------------
12). "What is the rationale for the proposed extensions later in the
paper, i.e. what are the use cases or evaluation-driven itersations?"

----------------------------------------------------------------------
Reviewer #3:
----------------------------------------------------------------------

----------------------------------------------------------------------
13). "Possibly the authors could signal at the end of section 1
something like "Readers unfamiliar with logic notation may safely
ignore the small amount they may encounter in what follows.""

[josh]: I like the reviewer's idea about a sentence at the end of
section 1 saying that "Readers unfamiliar with logic notation may
safely ignore the small amount they may encounter in what follows" and
focus on the examples accompanying each block description.

[mark]: I agree

----------------------------------------------------------------------
14). "Although not enough to recommend required additions, I am left a
little puzzled as to why the authors seem to be silent on two issues:
OWL data type properties and owlifier instance entities. Both of these
would seem easy to represent in their owlifier syntax. I would also
imagine that both could arise in representing and reasoning about both
dataset metadata and record-level metadata on biodiversity
databases. So I am left wondering whether I am wrong about the ease of
representing them in spreadsheets and attaching DL axioms, or whether
instead, the authors believe that these two OWL constructs are better
treated in as data for reasoners or elsewhere."

[josh]: I think we might need to be clearer that we are talking about
class / term structuring and not instantiating with data (class
instantiating comes later, after the formal conversion to OWL and by
the knowledge engineer).  Maybe this could be added to the figure in
some way.

I admit data-type properties are an obvious exclusion and should be
justified.  This could be done by simply stating that owlifier is a
mechanism for structuring classes with no additional, complicating
information at this stage (although we do allow for cardinality...).

[mark]: Yes, it is hard to convey the reason for this focus on
Classes.  Well, I feel that our use case is driven by using these
ontologies to annotate data-- values of which become the
instances. Maybe this is not a proper understanding.  But also even
Rector et al. OWL Pizzas paper states: "this paper only concerns
issues in defining OWL classes, since this is the strength of OWL-DL,
and most existing classifiers deal with individuals completely or not
at all." So we are in good company by just advancing capability to
work with classes and relationships.

Would such support for data-type properties be low-hanging fruit for
extending owlifier?  Especially the ability to capture enumerated
lists might be very useful in the future.

----------------------------------------------------------------------
15). "On page 4 (Subsection "Overlap blocks", the second sentence has a
small English ambiguity: "[...]a given set of classes may have
overlapping instances." could reasonably be understood to mean that
some instances overlap one another, whereas I believe the clause means
"several classes may have common instances." [Grammatically, the
overlapping instances do not belong to the collective "set of classes"
but rather to each of the members of that collective]"

[josh]: I agree with the grammatical comment to help the lay
reader. overlapping --> common

[mark]: ow about "Overlap blocks explicitly relax this assumption by
stating that a given set of classes may share common instances" (2nd
sent in para on Overlap blocks)

----------------------------------------------------------------------
16). "Reference 22 is missing 'c' in the article title.  "logi" should
be "logic"."

FIXED