### 2010-01-22 docs/algdraft/Materialize_DB.tex
Updated materialization algorithm.

\documentclass[10pt]{article}
%\title{}
%\author{}
%\date{}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{a4}
\usepackage{color}
\renewcommand\floatpagefraction{0.99}
\renewcommand\topfraction{0.99}
\renewcommand\bottomfraction{0.99}
\renewcommand\textfraction{.05}
\setcounter{totalnumber}{5}

%\setlength{\parindent}{0pt}
\setlength\topmargin{0in}
\setlength\headheight{0in}
\setlength\headsep{0in}
\setlength\textheight{8.5in}
\setlength\textwidth{6.5in}
\setlength\oddsidemargin{0in}
\setlength\evensidemargin{0in}

\newtheorem{example}{Example}[section]
\begin{document}

\section{Valid annotation constraints}
Among the {\em key}, {\em distinct}, and {\em identifying} constraints,
there are correlations between ``key yes'' and ``distinct yes''.
\begin{itemize}
\item ``Distinct yes'' on an observation type implies that its measurements collectively form a key of this observation type.
\item Logically, we can specify ``Distinct no'' on an observation type and specify ``key yes'' on all of its measurements.
In this case, when two rows (for the observation type) have the same values for these key measurements, they represent the different observation instances. It means that there may be other ``variables'' that can distinguish one observation instance from the other, but this dataset does not catch this variable ({\bf HP: is ``variable'' the right word to use? }). So, in real application, we do not encourage such situation to appear. 
\end{itemize}

An example with invalid annotation: given the data in Table \ref{tb:pltarea}. The following annotation is {\em invalid}.
According to the annotation, if two plots have the same value on EntityName, they represent the same plot observation. 
Obviously, it will have probelem to interprete the data in Table \ref{tb:pltarea} with this annotatioin. 
E.g., the first and second row catching plot with EntityName $A$ should be the same plot obervation.
However, the data shows that this one plot has two different areas $1.0$ and $1.1$, so, there is confusion here.
\\

\noindent
{\bf observation} ``o1''  \textcolor{blue}{distinct yes}\\
\verb|    |{\bf  entity} ``Plot''\\
\verb|    |{\bf measurement} ``m1'' \textcolor{blue}{key yes}\\
\verb|        | {\bf characteristic} ``EntityName'' \\
\verb|        | {\bf standard} ``Nominal''\\
\verb|    | {\bf measurement} ``m2''\\
\verb|        | {\bf characteristic} ``area'' \\
\verb|        | {\bf standard} ``sqft''\\
\verb|       |$\cdots$ Not finished....

\begin{table}[htb]
\begin{center}
%\begin{tabular}{cc}
%\begin{tabular}{|l|l|l|}
%\hline
%plt & spp & dbh\\\hline
%A & piru & 35.8 \\\hline
%A & piru & 36.2 \\\hline
%B & piru &33.2 \\\hline
%B&abba&34\\\hline
%\end{tabular}
%&
\begin{tabular}{|l|l|l|l|}
\hline
plt & area & spp & dbh\\\hline
A & 1.0 & piru & 35.8 \\\hline
A & 1.1 & piru & 36.2 \\\hline
B & 2.0 &piru &33.2 \\\hline
B& 2.0 & abba&34\\\hline
\end{tabular}\\
%(a) & (b)
%\end{tabular}
\end{center}
\vspace{-0.2in}
\caption{Dataset}
\label{tb:pltarea}
\end{table}

In our algorithm, we assume that the annotations on the dataset are valid, i.e., they comply with the correlation rules among these constraints.

{\bf From HP:} More to come on this part...

\section{Data materialization algorithm}
The {\bf MaterializeDB} algorithm catches the {\em key}, {\em distinct} and {\em identifying} constraints in the annotation during the materialization process. 
The input of the algorithm is the $Dataset$ and the annotations $A.*$ on it. 
Each row in the input dataset represents the information related to one or more observations and their contexts.
The input $A.*$ represents the annotation information.
More specifically,
\begin{itemize}
\item A.MeasType = $\{\langle \underline{MeasTypeId}, ObsTypeId, CharType, StdType, ProtType, Precision, isKey \rangle\}$;
\item A.ObservationType = $\{\langle \underline{ObsTypeId}, EntTypeId, isDistinct \rangle\}$ \\
	Note that, we hide the ``AnnotId'' in this schema, which denotes the resource the annotation is on.
    We do not include this in the algorithm because the algorithm focuses on dealing with annotations on one resource.
    This way, we can simplify the description.
\item A.ContextType = $\{\langle \underline{ObsTypeId, ContextObsTypeId, RelType}, isIdentify\rangle\}$
\item A.Map = $\{\langle \underline{MeasTypeId, ResAttribute, Cond}, Val\rangle\}$
\end{itemize}

The output of the algorithm is a set of materialized tables represented in the OBOE model and denoted by $OBOE.*$. In detail,
\begin{itemize}
\item $OBOE.Observation=\{\langle \underline{ObsId}, EntId \rangle\}$ keeps all the observation instances materialized from $Dataset$.
\item $OBOE.Measurement=\{\langle \underline{MeasId}, ObsId, MeasType, Val\rangle\}$
\item $OBOE.Entity=\{\langle \underline{EntId}, EntType\rangle\}$
\item $OBOE.Context =\{\langle \underline{ObsId, ConextObsId, ContextType}\rangle\}$
\end{itemize}

Algorithm \ref{alg:matdb} shows the framework of our algorithm. 
In this algorithm, $ObsIdx$ is the index structure maintained for the observation instances 
whose types are specified with {\em distinct yes}. 
This index maintains the mapping from observation type and key values to its corresponding observation instance id. 
Obviously, only when observation type is specified with {\em distinct yes}, i.e., we want to keep track of the same observation instances, we need to maintain their instances in this index. 
The key values can be calculated in three different cases.
In the first case where just one measurement of an observation type is specified with ``key yes'' the key value is the value of this ``key'' measurement. In the second case, several measurement types are marked with ``key yes'', the key value is the combined value of these several measurement. In the third case, this object type is marked with ``identifying yes'', the key value is the combined value of its context observation's key measurement values.

Similarly, in case that some measurement type(s) of an observation type is/are specified with ``key yes'', 
if different observation instances have the same value on these key measurements, semantically, we interpret that they are of the same entity. $Entidx$ is the index structure for tracking the distinct entities. 
The key value is computed in the same way as that for an observation instance.

The procedure of this algorithm is very straightforward. 
It processes the dataset in a row-wise manner. 
In dealing with each row, five steps are involved. 
The first step generates orphan measurement instances which are not connected any observation instances.
The second step group these measurement instances according to their observation types.
Then, for different observation types, we materialize entity instances and observation instances by either creating a new one or return an existing one. The last step assigns the context relationship among the different observation instances.

\begin{algorithm} [htb]
\caption{{\bf MaterializeDB} ($Dataset, A.*$)}
\label{alg:matdb}
{\small
\begin{algorithmic}
\STATE /* {\em Dataset}: [Input] in the form of a flat file */\\
\STATE /* {\em A }:  [Input] Annotations*/
\STATE
\STATE $ObsIdx =\emptyset$;  /* Keep index $\langle ObsTypeId, KeyVal\rangle \rightarrow ObsId$*/
\STATE $EntIdx = \emptyset$;  /* Keep index $\langle ObsTypeId, KeyVal\rangle \rightarrow EntId$*/
\STATE
\FOR { (each $Row\langle A1, A2, \cdots, An\rangle \in Dataset$)}
    \STATE /* Step 1: Define measurement instances */\\
            $MeasSet$ = CrtMeasurement($Row, A.*$);
    \STATE
    \STATE /* Step 2: Partition the measurement instances according to observation types*/\\
        $ObsType2MeasIdx$ = PartMeas($MeasSet, A.*$);
    \STATE
    \STATE $ContextIdx = \emptyset$;  /* Keep index $ObsTypeId \rightarrow ObsId$ to materialize context*/
    \FOR{(each $ObsTypeId \in ObsType2MeasIdx.keys$)}
      \STATE /* Step 3: Find or create the entity instance for each observation type partition */
        \STATE $EntId$ = MaterializeEntity($ObsTypeId, ObsType2MeasIdx, EntIdx, A.*, OBOE.*$);
        \STATE
        \STATE /* Step 4 Find or create the observation instance for each observation type partition */
        \STATE MaterializeObs($ObsTypeId, EntId, ObsType2MeasIdx, ObsIdx, ContextIdx, OBOE.*$);
    \ENDFOR
    \STATE
    \STATE /* Step 5: Assign the context observation instances */  \\
            MaterializeContext($ContextIdx, A.*, OBOE.*$);
\ENDFOR
\RETURN OBOE;
\end{algorithmic}
}
\end{algorithm}

{\bf Analysis of MaterializeDB}. \\
{\bf Time}: As can be seen that  this algorihm scans the original data file in a row-by-row manner without revisiting the already seen rows. So, it is linear in the size of the dataset. 
 We use $EntIdx$ and $ObsIdx$ to facilitate the checking of unique entity and observation instances. Let $m$ be the number of unique keys, each checking could take O(log(m)) time. 
So, in total, the algorithm runs in O(n log(m)) where $m\ll n$ generally. \\
{\bf Space}: $EntIdx$ and $ObsIdx$ are the intermdedite structures that we use in the algorithm. Since they keep the unique entity and observation key values. The space complexity is of $O(m)$.

\begin{algorithm} [htb]
\caption{{\bf CrtMeasurement} ($Row, A.*$)}
{\small
\begin{algorithmic}
    \STATE/* Create new orphan measurement instances*/
    \STATE $MeasSet = \emptyset$; /* Keep the set of new measurement instances*/
	\FOR {(each $m = \langle MeasTypeId, ResAttribute, Cond, Val\rangle \in A.Map$)}

		\STATE {\bf if}{(($m.ResAttribute != Row.Ai.Attrname$) OR ($Row.Ai$ does not satisfy $m.Cond$))}
		  {\bf continue};
		%\STATE $ObsTypeId$ = GetObsTypeId($A.MeasType, m.MeasTypeId$);
		%\STATE
		\STATE $mi_{id}$ = GetNewMeasId($OBOE.Measurement$);
		\STATE {\bf if} ($m.Val!=NULL$) $MeasVal$ = $m.Val$;
		\STATE {\bf else} $MeasVal$ = $Row.Ai.Val$;
		\STATE Create a measurement instance $\langle mi_{id}, null, MeasType, MeasVal\rangle$ and add it to $MeasSet$;
	\ENDFOR
    \RETURN $MeasSet$
\end{algorithmic}
}
\end{algorithm}

\begin{algorithm} [htb]
\caption{{\bf PartMeas} ($MeasSet, A.*$)}
{\small
\begin{algorithmic}
    \STATE /* Partition measurement instances according to their observation types*/
    \STATE $ObsType2MeasIdx=\emptyset$ /* Keep index for $ObsTypeId \rightarrow \{mi\}$ */
    \FOR{(each $mi \in MeasSet$)}
    \STATE $ObsTypeId$ = GetObsTypeId ($A.MeasType, mi.MeasTypeId$);
	\STATE Update $ObsType2MeasIdx$ by changing the item $ObsTypeId \rightarrow \{mi\}$;
    \ENDFOR
    \RETURN $ObsType2MeasIdx$;
\end{algorithmic}
}
\end{algorithm}

\begin{algorithm} [htb]
\caption{{\bf MaterializeEntity}($ObsTypeId, ObsType2MeasIdx, EntIdx, A.*, OBOE.*)$}
{\small
\begin{algorithmic}
\STATE $KeyVal$ = GetObsTypeKeys ($ObsTypeId, ObsType2MeasIdx$);
\STATE $HasKey$ = false;
\STATE {\bf if }{($ObsTypeId$ has key measurements OR is specified with distinct yes)}
     $HasKey$ = true;
\STATE $EntType$ = GetObsEntityType ($A.ObservationType, ObsTypeId$)
\STATE $CrtNewEntInst$ = true;
\IF{ ($HasKey$==true) }
	\STATE $EntId$ = GetEntId($ObsTypeId, KeyVal, EntIdx$); %/*From the current entity key index get the id of the entity instance*/
	\STATE {\bf  if} {($EntId!=NULL$)} $CrtNewEntInst$ = false;
\ENDIF
\IF{($CrtNewEntInst==true$)}
	\STATE $EntId$ = CrtEntId($EntType$);
	\STATE Create an entity instance $ei = \langle EntId, EntType\rangle $ and put $ei$ to $OBOE.Entity$;
	\STATE {\bf if} ($HasKey$==true) $EntIdx = EntIdx \cup \{\langle ObsTypeId, KeyVal \rangle \rightarrow EntId$\};
\ENDIF
\RETURN $EntId$;
\end{algorithmic}
}
\end{algorithm}

\begin{algorithm} [htb]
\caption{{\bf MaterializeObs}($ObsTypeId, EntId, ObsType2MeasIdx, ObsIdx, ContextIdx, OBOE.*$)}
{\small 
\begin{algorithmic}
    \STATE $KeyVal$ = GetObsTypeKeys ($ObsTypeId, ObsType2MeasIdx$);
    \STATE $IsObsDistinct$ = CheckIfObsDistinct($A.ObservationType, ObsTypeId$);
		\STATE $CrtNewObsInst$ = true;
	\IF{ ($IsObsDistinct$==true) }
        %/*From the current observation key index get the id of the observation instance*/
		\STATE $ObsId$ = GetObsId($ObsTypeId, KeyVal, ObsIdx$);
		\STATE {\bf if}{($ObsId!=NULL$)}  $CrtNewObsInst$ = false;
	\ENDIF
	\IF{($CrtNewObsInst==true)$}
		\STATE Create an observation instance $oi = \langle ObsId,  EntId\rangle$ and put $oi$ to $OBOE.Observation$;
		\STATE {\bf  if} ($IsObsDistinct$==true) $ObsIdx = ObsIdx \cup \{\langle ObsTypeId, KeyVal \rangle \rightarrow ObsId$\};
           \ENDIF
    \STATE
    \STATE /*Maintain the measurement instances for this observation*/
    \STATE $miSet$ = GetMeasInst($ObsType2MeasIdx, ObsTypeId$);
    \STATE {\bf if} {($ObsId$ is a new one)}
         Set the $obsId$ to each $mi \in miSet$ so that $mi$-s are not orphans;
    \STATE {\bf else} 
         Discard all the $mi \in miSet$;

    \STATE Put all the $mi \in miSet$ to $OBOE.Measurement$;
    \STATE

	\STATE $ContextIdx = ContextIdx \cup \{ObsTypeId \rightarrow ObsId\}$; /* $ContextIdx$ is also output*/

\end{algorithmic}
}
\end{algorithm}

\begin{algorithm} [htb]
\caption{{\bf MaterializeContext}($ContextIdx, A.*, OBOE.*$)}
{\small 
\begin{algorithmic}
\FOR{ ($ObsTypeId \rightarrow ObsId \in ContextIdx$)}
\STATE $ContextObsTypeId, Rel$ = GetContextObsTypeRel($A.ContextType, ObsTypeId$);
\IF{($ContextObsTypeId!= NULL$)}
	\STATE $ContextObsId$ = GetContextObsId($ContextIdx, ContextObsTypeId$);
	\STATE Create a context instance $ci=\langle ObsId, ContextObsId, Rel\rangle$;
	\STATE Put $ci$ to $OBOE.Context$;
\ENDIF
\ENDFOR
\end{algorithmic}
}
\end{algorithm}

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|l|l|l|}
\hline
yr & spec & spp & dbh\\\hline
2007 & 1 & piru & 35.8 \\\hline
2008 & 1 & piru & 36.2 \\\hline
2008 & 2 & abba & 33.2 \\\hline
\end{tabular}
\end{center}
\vspace{-0.2in}
\caption{Dataset 1}
\label{tb:dataset1}
\end{table}

\begin{example} [Example with ``key yes'' and ``distinct yes'', without ``identifying yes'']
Take the data in Table \ref{tb:dataset1} \footnote{This is in page 6 in the powerpoint file.}
as an example to explain the algorithm. \\
For {\bf Row(2007, 1, piru, 35.8)}
\begin{itemize}
\item Step 1 creates four measurement instances: 
$\langle mi_1, null, Year, 2007\rangle$, $\langle mi_2, null, DBH, 35.8\rangle$, \\
$\langle mi_3, null, TaxonomicTypeName, Picea~rubens\rangle$, $\langle mi_4, null, EntityName, 1\rangle$,\\
and returns $MeasSet = \{mi_1, mi_2, mi_3, mi_4\}$;
\footnote{For all the instances, the measurement characteristic is set to represent Measurement Type}.

\item Step 2 returns $ObsType2MeasIdx = \{\{o_1\rightarrow \{mi_1\}, o_2 \rightarrow \{mi_2, mi_3, mi_4\}\}$. 
 \item Step 3-4: for each observation types $o_1$ and $o_2$,  materialize entity and observation instance
    \begin{itemize}
    \item for $o_1$ (with associated instance $mi_1$ of type $m_1$
    \begin{itemize}
        \item Since $m_1$ is specified as key, get the $KeyVal=2007$;
        \item No entity with this key exists in $EntIdx$, create an entity $\langle ei_1, TemporalRange\rangle$; Now, $EntIdx=\{\langle o_1, 2007\rangle\rightarrow ei_1\}$.
        \item Since $o_1$ is specified as {\em distinct}, need to make sure we do not create redundant observation instances. No entry with the key $\langle o_1, 2007\rangle$ exists in $ObsIdx$, so, create an observation instance $oi_1$, which is of entity $ei_1$ and represented as $\langle oi_1, ei_1\rangle$. \\
	Now, $ObsIdx=\{\langle o_1, 2007 \rangle \rightarrow oi_1\}$
        \item Connect $mi_1$ to $oi_1$;
    \end{itemize}

    \item When deal with $o_2$,
        \begin{itemize}
        \item $KeyVal=1$.
        \item Create an entity instance $\langle ei_2, Tree\rangle$; $EntIdx=\{\langle o_1, 2007\rangle\rightarrow ei_1, \langle o_2, 1\rangle\rightarrow ei_2\}$.
        \item Create an observation $\langle oi_2, ei_2 \rangle$.
            No need to update $ObsIdx$ because $o_2$ is not identified as {\em distinct}.
        \item Connect $mi_2$, $mi_3$ and $mi_4$ to $oi_2$;
        \end{itemize}
    \end{itemize}
    \item Step 5 assigns the context relationship between $oi_1$ and $oi_2$;
    \end{itemize}

For {\bf Row (2008, 1, piru, 36.2)}
    \begin{itemize}
    \item Step 1 creates measurement instances $\langle mi_5, null, Year, 2007\rangle$,
    $\langle mi_6, null, DBH, 35.8\rangle$,\\
    $\langle mi_7, null, TaxonomicTypeName, Picea~rubens\rangle$,
    $\langle mi_8, null, EntityName, 1\rangle$\\
    and returns $MeasSet = \{mi_5, mi_6, mi_7, mi_8\}$;
    \item Step 2 gets $ObsType2MeasIdx = \{\{o_1\rightarrow \{mi_5\}, o_2 \rightarrow \{mi_6, mi_7, mi_8\}\}$
    \item Step 3-4: for each observation types $o_1$ and $o_2$ materialize entity and observation instance
        \begin{itemize}
            \item for $o_1$
            \begin{itemize}
            \item  $KeyVal=2008$;
            \item Create an entity instance $\langle ei_3, TemporalRange\rangle$; \\
                $EntIdx=\{\langle o_1, 2007 \rangle \rightarrow ei_1, \langle o_2, 1 \rangle \rightarrow ei_2, \langle o_1, 2008 \rangle \rightarrow ei_3\}$.
             \item Create an observation instance $\langle oi_3, ei_3 \rangle$;\\
              $ObsIdx=\{\langle o_1, 2007\rangle\rightarrow oi_1, \langle o_2, 1\rangle\rightarrow oi_2, \langle o_1, 2008\rangle\rightarrow oi_3\}$
		\item Connect $mi_5$ to $oi_3$; 
            \end{itemize}
            \item When deal with $o_2$,
            \begin{itemize}
             \item $KeyVal=1$.
            \item {\bf item $\langle o_2, 1 \rangle \rightarrow ei_2$ is already in $EntIdx$, so get the entity id $ei_2$. No need to create an entity.}
            \item Since $o_2$ is not specified with {\em distinct yet}, we NEED to create an observation $\langle oi_4, ei_2 \rangle$. No need to update $ObsIdx$.
		\item Connect $mi_6, mi_7, mi_8$ to $oi_4$; 
            \end{itemize}
        \end{itemize}
    \end{itemize}
For {\bf ROW (2008, 2, abba, 33.2)}
    \begin{itemize}
        \item For $o_1$'s measurement $2008$, 
	\begin{itemize}
	\item Since $\langle o_1, 2008 \rangle \rightarrow ei_3$ already exists in $EntIdx$, {\bf no need to create a new entity}.
        \item Since $o_1$ is specified with {\em distinct yes}, and $\langle o_1, 2008 \rangle \rightarrow oi_3 $ already exists in $ObsIdx$, 
	{\bf no need to create a new OBSERVATION} and no need to put the measurement instance for $2008$ into OBOE model.
	\end{itemize}
    \end{itemize}

\end{example}

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|l|l|}
\hline
plt & spp & dbh\\\hline
A & piru & 35.8 \\\hline
A & piru & 36.2 \\\hline
B & piru &33.2 \\\hline
\end{tabular}
\end{center}
\vspace{-0.2in}
\caption{Dataset 2}
\label{tb2}
\end{table}

\begin{example} [Example with identifying]\label{eg2}
Let us use the data in Table \ref{tb2} as an example.
Before we go through the algorithm step by step,  we first note that  $o_1$ and $o_2$ have key measurements $m1$ and $m_2$ respectively. 
So, we need to maintain the distinct entity instances for both of these two observation types. 
In addition, $o_1$ is specified with {\em distinct yes} while $o_2$ is not. So, we need to maintain the distinct observation instances for $o_1$ but not for $o_2$.

For the {\bf first row}, 
\begin{itemize}
\item The first step generates three measurement instances $MeasSet=\{\langle mi_1, null, EntityName, A\rangle$, \\
$\langle mi_2, null, TaxonomicTypeName, Picea~rubens\rangle$, $\langle mi_3, null, DBH, 35.8\rangle\}$.
\item The second step gets $ObsType2MeasIdx = \{\{o_1\rightarrow \{mi_1\}, o_2 \rightarrow \{mi_2, mi_3\}\}$.
\item For each observation type, create entity and observation instances.
\begin{itemize}
\item For $o_1$, the key value is $A$. Since there is no such a key in $EntIdx$, we create an entity $ei_1$ of type $Plot$.
	$EntIdx = \{\langle o_1, A\rangle \rightarrow ei_1\}$. \\
	We create an observation instance $oi_1$ whose entity is $ei_1$.\\
	$ObsIdx = \{\langle o_1, A\rangle \rightarrow oi_1\}$.
	Connect the measurement instances to observation instances  $mi_1 \rightarrow oi_1$,
\item For $o_2$, 
	the key value is $(A, Picea~rubens)$ since it has context $o_1$ with ``{\em identifying yes}'',  we create an entity instance $ei_2$ of type $Tree$. \\
	$EntIdx = \{\langle o_1, A\rangle \rightarrow ei_1, \langle o_2, (A, Picea~rubens) \rangle \rightarrow ei_2\}$.\\
	We  create an observation instance $oi_2$ whose entity is $ei_2$.\\
	Connect the measurement instances $mi_2$ and $mi_3$  to observation instance $oi_2$.
\end{itemize}
\item The last step for this row is to connect the observations using context relationship. For this instance, we connect $oi_1$ to $oi_2$ with context ``{\em Within}''.
\end{itemize}

For the {\bf second row}, 
\begin{itemize} 
\item The first step defines three measurement instances $MeasSet=\{\langle mi_4, null, EntityName, A\rangle$, \\
$\langle mi_5, null,TaxonomicTypeName, Picea~rubens\rangle, \langle mi_6, null, DBH, 36.2\rangle\}$.
\item The second step gets $ObsType2MeasIdx = \{\{o_1\rightarrow \{mi_4\}, o_2 \rightarrow \{mi_5, mi_6\}\}$.
\item For each observation type, create entity and observation instances.
\begin{itemize}
\item For $o_1$, the key value is $\langle o_1, A\rangle$, $EntIdx$ already has an item for it with entity instance $ei_1$. No need to create a new instance for it. \\
To create observation instance, since $o_1$ is specified with ``{\em distinct yes}'' and the key value is $\langle o_1,A\rangle$, 
which corresponds to observatin instance $oi_1$. So, we do not need to create a new observation for it.\\
When we try to connect the measurement instance $mi_4$ to observation instance,  we realize that we did not create a new observation instance for type $o_1$. 
So its related measurement instance $mi_4$ can be discarded. 
\item For $o_2$, the new key value is $\langle o_2, (A, Picea~rubens)\rangle$ , which corresponds to $ei_2$ in $EntIdx$, so no need to create a new instance for it either. \\
To create observation instances, since no ``{\em distinct yes}'' is specified, we create a new observation instance $oi_3$ for it.
Then, we connct the measurement instances $\{m_5, m_6\}$ to observation instance $oi_3$.
\end{itemize}
\end{itemize}

When we process the {\bf third row}, we have a new key value $\langle o_1, B\rangle$ for $o_1$, thus we create a new entity instance for it.
For $o_2$, we have new key value $\langle o_2, (B,~Picea~rubens)\rangle$ and create a new entity instance for it.

\end{example}

\end{document}