### 2010-05-07 docs/algdraft/DataIntegration/DataIntegration.tex
Refine the definition and add introduction. Is refining the materialization part.

\documentclass[conference]{IEEEtran}

\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

\usepackage{algorithm}
\usepackage{algorithmic}
 \renewcommand\floatpagefraction{0.99}
 \renewcommand\topfraction{0.99}
 \renewcommand\bottomfraction{0.99}
 \renewcommand\textfraction{.05}
 \setcounter{totalnumber}{5}

\newtheorem{example}{Example}[section]
\newtheorem{definition}{Definition}[section]
\newcommand{\from}[2]{{\bf[{\sc from #1:} #2]}}

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{Querying Integrated Scientific Observation and Measurement Data}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Michael Shell}
\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
Georgia Institute of Technology\\
Atlanta, Georgia 30332--0250\\
Email: http://www.michaelshell.org/contact.html}
\and
\IEEEauthorblockN{Homer Simpson}
\IEEEauthorblockA{Twentieth Century Fox\\
Springfield, USA\\
Email: homer@thesimpsons.com}
\and
\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
\IEEEauthorblockA{Starfleet Academy\\
San Francisco, California 96678-2391\\
Telephone: (800) 555--1212\\
Fax: (888) 555--1212}}

\maketitle

\begin{abstract}
%\boldmath
The abstract goes here.
\end{abstract}

% no keywords

\section{Introduction}
In this work, we study the query processing over scientific
observation and measurement data using OBOE
model\cite{DBLP:conf/er/BowersMS08}. OBOE model is a conceptual model
used to interpret observation and measurement data.

\subsection{Background}

%Application background
In many scientific domains (e.g., ecology, hydrology, earth science,
geology), people collect observational data. Such data
record the observed value of some real world entity at some specific
place and time. E.g., ecologists studying relationship between the 
growth pattern and the treatments often need to record the tree
heights. The collected data reflect the fact the tree height of a
specific tree is 30.1in on May 1, 2009 and 30.3in on May 1, 2010.

%Characteristics of scientific observational data: non-normalized
Almost all such scientific data do not follow database higher normal
forms. Generally, scientists have their way in interpreting their
data, but they are not ready for any normalization process. 
For example, Table \ref{tb:dataset} is a simplified but typical dataset collected by
a scientist who study the growth pattern of trees. 
Obviously, the ``plt'' of the first two rows is the same place, and
the ``plt'' of the last two rows is the same. 
Here, their area information is redundant. 
In the real application, many columns have redundant information.

%HP: this one has the coordinates, it's troublesome to explain. 
%% \begin{table}[htb]
%% \begin{center}
%% \begin{tabular}{|l|l|l|l|l|}
%% \hline
%% code & no & ht & plt & area\\\hline
%% piru & 1 & 35.8 & $(34^{\circ}8'3N,118^{\circ}14'37W)$ & 4.0\\\hline
%% piru & 2 & 36.2 & $(34^{\circ}8'3N,118^{\circ}14'37W)$ & 4.0\\\hline
%% piru & 1 & 25.7 & $(36^{\circ}10'30N,115^{\circ}8'11W)$ & 3.0 \\\hline
%% %abba & 1 & 15.6 & $(36^{\circ}10'30N,115^{\circ}8'11W)$ & 3.0\\\hline
%% capo & 1 & 15.6 & $(36^{\circ}10'30N,115^{\circ}8'11W)$ & 3.0\\\hline
%% \end{tabular}
%% \end{center}
%% \caption{Dataset}
%% \label{tb:dataset}
%% \end{table}

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|l|l|l|l|}
\hline
code & no & ht & plt & area\\\hline
piru & 1 & 35.8 & California & 4.0\\\hline
piru & 2 & 36.2 & California & 4.0\\\hline
piru & 1 & 25.7 & Oregon & 3.0 \\\hline
%capo & 1 & 7.8 & Oregon & 3.0\\\hline
Oriental poppy& 1 & 7.8 & Oregon & 3.0\\\hline 
\end{tabular}
\end{center}
\caption{Dataset}
\label{tb:dataset}
\end{table}
%HP: California poppy info
%Oregon iris: http://en.wikipedia.org/wiki/List_of_native_Oregon_plants
%http://en.wikipedia.org/wiki/California_poppy

%Introduce major concepts
In scientific domains where people collect observation and measurement
data, there are several commonly used and widely recognized canonical
concepts(\cite{oboe, om}). 
In this paper, we refer these concepts as OM concepts. 
The canonical concepts include {\em observation}, {\em measurement},
{\em characteristic}, {\em standard}, {\em protocol} or ({\em procedure}). 
For example, {\bf add an example to illustrate these concepts. }

%Introduce annotation, Motivation to use annotation
Generally, every scientific dataset goes into the data repository with
some metadata, e.g., Dublin core metadata\cite{***}, Darwin core metadata\cite{***}. 
However, all these metadata are at the dataset level, they did not
provide enough information for the data {\em content} inside the
dataset. 
To better make use of the data content in the repository, more systems
are embracing the ideas of having metadata on the data content. 
E.g., Some systems \cite{tdar} provides a mechanism to collect the
column/attribute level metadata, some system \cite{semtools} uses
{\em annotation} to add more semantic information to the data
content. In this work, we follow the terminology in \cite{semtools} and
use the term {\em annotation} to distinguish the 
metadata on the data content from those on data objects at a coarser
granularity. 
%How to represent the annotation and the symbols used
We propose tools for scientists to provide annotations to scientific
observational data. 
\from{HP}{Put a screen-dump of annotation.}

%Challenge to query such data
\from{HP}{Queries to such data and challenges to answer such queries.}
When a user formulates a query, it is unrealistic for a him/her to know the 
underlying data structure of the data.  
But they all know the well recognized OM concepts as discussed above. 
So, naturally, when searching such scientific data, people are more
interested in finding datasets related to such OM terminologies as observation
measurement, characteristic, etc. 
Given the dataset in Table \ref{tb:dataset}. 
People may ask the following queries.

\begin{example}\label{eg:query}
Simple and summarization queries: 
\begin{itemize}
\item $Q_1$: Give me the data sets that contain species ``Picea rubens'' observations.
%$Q_2$: Give me the data sets that have measurements on ``area''  characteristics. 
\item $Q_2$: Give me the data sets that contain species ``Picea rubens''
  observations in ``California''. 
\item $Q_3$: Give me the data sets that contain at least five distinct
  ``Picea rubens'' observations.
\item $Q_4$: Give me the data sets that have trees with average ``height''
  than 20.0 in ``California''. 
% Shawn: an example on summarize one measure with respect to another.
% This could take two forms: one where the context forms the groups
% (like by year), and another where the two measurements are within the
% same context (like finding correlations).
\end{itemize}
\end{example}

\from{HP}{Current data integration effort}

\from{HP}{Our effort in querying observational data: introduce OBOE, query}

\from{HP}{Need re-organization. The description in the
  following several sections may be moved to here later.}

\subsection{Contribution and paper organization}
Contributions of this work:
\begin{itemize}
\item Formalize the queries over observation and measurement query.
\item Propose three methods to evaluate queries. 
\end{itemize}

This paper is organized as follows. 
Section \ref{sec:relatedwork} reviews works that are related to this
research. Section \ref{sec:dataquery} formalize the data model and the
queries that people are interested to ask.

\section{Related work}\label{sec:relatedwork}

\from{HP}{This will come after the real problem definition and the
  solution. The description in the following several sections may be
  moved to here later.}

\section{Data model, annotation and query}\label{sec:dataquery}

In this section, we first illustrate the data model. Then, we
formalize the queries that scientists in this domain tend to ask.

\subsection{Data model}\label{sec:datamodel}

%Data model: Data, attribute, data value set
When a scientist contributes data into an integrated data repository,
a widely accepted way is to convert each dataset to a data
table\cite{tdar} or treated as a separate object entity\cite{knb}. 
{\bf (HP: is it really widely used? any other system uses this
  way?. Add more citations here.)}
In using this method, the database contains metadata of each
dataset and the definition of the data table (e.g., attribute/column
name, attribute type, etc).

%HP: 
%An alternative way to organize the data is to store them in one big
%data table with three columns data\_file, attribute\_name,
%attribute\_value. The problem with this structure is that the
%attribute\_value are different types.
In our work, we focus on querying this kind of data repository (or
databases). To formalize the scenario, we use $D$ to denote the set of data
tables in the data repository and $d$ to refer to a specific data
table. Each data table $d$ contains metadata about the attribute
definition $Attr_{d}$. Sometimes, we also use {\em column} to refer to an
attribute. 
Given one data table $d$, an attribute $attr_i \in Attr_{d}$ or column index $i$, 
$d[attr_i]$ or $d[i]$ represents the set of values for the attribute
$attr_i$ or for the $i$-th column.
%HP: Implementation detail (no need to put it in the paper): 
%When we put the data to data tables, we add a unique record id for
%each row in the data. This is used to get the unique record id.

\subsection{Annotation}\label{sec:annotation}

%What kinds of annotation information is needed.

We use $A$ to denote the annotation of one dataset. 
Internally, we keep the following information for the
annotation.

For data table $d$ with annotation $A$, the system keeps the
information $AD$ = $\{(d_{id}, A_{id}, d_{meta})\}$.
As we described before, we have four main objects to describe: 
observation type (OT), measurement type (MT), and
context type (CT), and the mapping (Map) from the measurement type to
resource attributes.  
So, the annotation contains the following information.

\begin{itemize}
\item $A.OT$ = $\{(\underline{A_{id},ot_{id}}, et, isDistinct)\}$ to
  describe an observation type, which denotes on which
  entity type (object in the real world) the observation is made. 
Very often, more than one observation can be made on one entity. Thus,{\em
  isDistinct} is used to denote whether the same value of key
meaasurements of an obbservation types can uniquely identify one
observation or not. 
\item $A.MT$ = $\{(\underline{A_{id}, mt_{id}}, ot_{id},isKey,
  Cha, \cdots)\}$. Generally, $MT$ contains information about
  characteristic ($Cha$), Standard, Protocol and Precisio. We do not include them
  here to make the description clearer. Here, {\em isKey} is used to denote
  whether one measurement type is the key measurement for the
  observation type $OT_{id}$ or not. 
\item $A.CT$ = $\{(\underline{A_{id}, ot_{id}, cot_{id},
  Rel}, isIdentify)\}$ where $cot$ represents the context observation
  type. 
\item $A.Map$ = $\{(\underline{A_{id},mt_{id}, attr_i, Cond},
  Val\}$
\end{itemize}

We use the following example to illustrate the concepts in the
annotation.

\begin{example}
$A.OT$ = \{($A_1, OT_1$, tree, false),($A_1, OT_2$, GeoSpot, true) \}
  denotes that annotation $A_1$ are for two observation types. One
  type $OT_1$ is on real world entity ``tree'' and the values of its key measurements
  do not uniquely identify a distinct observation. The other type
  $OT_2$ is for real world entity ``GeoSpot'' (i.e., geospatial
  location) and its key measurement types can identify it's unique
  observations. \from{HP}{Introduce key measurements before this.}

$A.MT$ =\{($A_1, MT_1, OT_1$, Species, true), 
($A_1, MT_2, OT_1$, SpecNo, true), 
($A_1, MT_3, OT_1$, Height, false),
($A_1, MT_4, OT_2$, Plot\_State, true), 
($A_1, MT_5, OT_2$, Plot\_Area, false)\} represents that $OT_1$
(for ``tree'') has three measurement types ($MT_1, MT_2$, $MT_3$) for characteristics
``Species'', ``SpecNo'', and ``Height''. The fist two measurement
  types together form the key measurement type for $OT_1$. 
$OT_2$ (for ``GeoSpot'') has one key measurement type $MT_4$ for
  characteristic ``Plot\_State'' and another measurement type $MT_5$
  for characteristic ``Plot\_Area''.

$A.CT$ = \{($A_1$,$OT_1$, $OT_2$, within, true)\}. It shows that
observation of a {\em tree} is within the context of a geo-spatial
location. And this context is used to identify the uniqueness of the
observation. E.g., a tree with species name ``piru'' and species no
$1$ in California
%$(34^{\circ}8'3N,118^{\circ}14'37W)$ (Los Angeles in California) 
is different from a tree with 
the same species name ``piru'' at Oregon. 
%location $36^{\circ}10'30N,115^{\circ}8'11W$ (Las Vegas in
%California).

%HP coordinates material: 
%http://www.traveljournals.net/explore/united_states/map/u1662328/los_angeles.html
%http://en.wikipedia.org/wiki/Las_Vegas,_Nevada

$A.Map$ = \{($A_1$, $MT_1$, ``code'', ``eq `piru''', ``Picea rubens''),
%($A_1$, $MT_1$, ``code'', ``eq `abba''', ``Abies balsamea''),
%($A_1$, $MT_1$, ``code'', ``eq `capo''', ``California poppy''),
($A_1$, $MT_2$,``no'', null,), 
($A_1$, $MT_3$, ``plt'', null,), 
($A_1$, $MT_4$,``area'', null,\}. 
The first mapping rules maps ``code''
attribute to the measurement type $MT_1$ (for ``Species''
characteristic) and change the value to ``Picea rubens'' if the code
is ``piru''. The meaning of the other mapping rules are very
obvious here, so we skip the details. 
\end{example}

\subsection{Observation and measurement (OM) query}
Above we defined the data model and the annotation, in this
sub-section, we define the queries that scientists are interested in
We denote the queries on the observational data as {\em Observation
and measurement (OM) query}.

In using the observational data, people are generally interested in
the following concepts: {\em observation, measurement,characteristic or standard}.

\begin{definition}[Basic OM query]\label{def:basic_omq}
A basic observation and measurement query is defined as $Q$::={\tt
  concept:cond}. 
Here, {\tt concept} is the main term in an observation model. 
{\tt cond} is in the form of $f([distinct]~attr)~op~value$ where $op$ is
basic operator from $\{=, \neq, >, <\}$ and $f$ is an
aggregation function from $\{sum, avg, count, min, max\}$. 
\end{definition}

\begin{definition}\label{def:qresult}
The result of a query $Q$ is a set of data objects (e.g., data tables) 
$\{d|d\in D \wedge d~satisfies~{\tt cond}\}$. 
For each of such result $d$, we use $d~s.t.~Q$ to denote that
data object $d$ satisfies the query. 
\end{definition}

Based on the different {\tt cond} definition, ``$d~satisfies~{\tt cond}$''
is translated into different formulas. 
E.g., if {\tt cond} is defined that the value of attribute ``area'' need to be
smaller than $3.0$, then, ``$d~satisfies~{\tt cond}$'' is translated to
``$dv<3.0|dv\in d[area]$''.

Definition \ref{def:basic_omq} can formulate queries on the
one specific observation and measurement. 
In real application, we need to consider the context relationship. 
So, we generalize the basic OM query to contextualized OM query as
follows.

\begin{definition}[Contextualized OM query]\label{def:context_omq}
$CQ$::=$Q_1~${\tt context}($Q_1.concept, Q_2.concept$)$~Q_2$. Here $Q_i$ is a basic query
in the form of {\tt concept:cond}. 
\end{definition}

For example, the queries in Example \ref{eg:query} can be formalized
as the following formal OM queries.

%\begin{example}
\begin{itemize}
\item $Observation: \langle Species='Picea~rubens'\rangle$
\item $Observation_1: \langle Species='Picea~rubens'\rangle$\\
  $IN(Observation_1, Observation_2$) \\
$Observation_2: \langle Meas_{unknown}=`California'\rangle$
\item $Observation: \langle Species='Picea~rubens' $\\
$\wedge
  count(distinct~Species)\geq 5\rangle$
\item $??$
\end{itemize}

%\end{example}

%\from{HP}{Add experiement: show the error rate (e.g., precision, recall, (or
%  directly $F_1$) of this method.? not needed?}

\section{Querying annotated and integrated OM data}

In this section, we discuss several methods that we can use to
answer OM queries over scientific observational data with
annotations.

%\subsection{Naive method}\label{sec:naive}

A very direct way search such data is to extract all the content in
the data cells and index them, this way, the system may help answer
very simple queries with the help of some thesaurus. 
For example, for $Q_1$ in Example \ref{eg:query}, as long as we can
translate the abbreviation ``piru'' in the dataset to ``Picea rubens''
based on some predefined thesaurus. We can find the results. 
Even for this kind of simple query, however, if we want to find
observation of ``California poppy'', an approximate match may return
the dataset in Table \ref{tb:dataset} as a result because it contains
both {\em California} and {\em Poppy}.
If such compound word search can be alleviated using some existing
techniques \cite{***}, %citation on compound word querying. 
it is almost impossible to directly use such IR-style system to answer a query
in the form of $Q_3$ and $Q_4$. 
I.e., answer queries with context.

This naive method cannot answer the desired query. It is because it
fails to catch the semantics in the data. 
In what follows, we would use the annotation during our search to get
more semantic support.

As we analyzed in Section \ref{sec:relatedwork}, there are two extrems
in querying integrated data. 
One is to rewrite the original query to a series of new queries over
the data; the other is to materialize the data to a consistent data
model and then answer the query over the materialized database. 
We what follows we work from these two directions to leverage the
semantic information in answering such queries.

\subsection{Query rewriting}\label{sec:queryrewrite}

As described in Section \ref{sec:dataquery}, 
an OM query is represented using 
OM concept terminologies such as observation
measurement, characteristic, etc. 
To answer such query from the original data model (i.e., dataset
metadata and data tables), we need to {\em rewrite} the quey over the
real data table. 
In what follows, we first sketch the re-writing process. Then, we
detailed the procedure in using the data model and
structures. Finally, we include the process for the more complicated
cases (e.g., with distinct, aggregate).

Roughly speaking, the query rewriting consists of two steps:
\begin{itemize}
\item From the given query, find the relevant data tables and
  attributes that need to be used answer the given query. 
\item Translate the given query to queries over the relevant data
  tables. 
\end{itemize}

{\bf TODO: put a figure here to illustrate the steps}.

The first step is to map the OM query to the real data structure. 
We can utilize the annotation structure $A.Map$ which keeps the correspondences between
measurement type and table attributes. 
So, when the given concept is {\em measurement}, we can directly get
its related attribute names and annotation id $A_{id}$. 
In case that the given OM concept is a non-measurement concept, we can use
other Annotation structures $A.*$ to figure out the measurementt types. 
E.g., if the OM concept is Characteristic ``Species'', we can use
$A.MeasType$ to get the measurement type id.

As mentioned in Section \ref{sec:annotation}, when some annotations are done over data tables, the system keeps the
information $\langle A_{id}, DTableId\rangle$. 
With the $A_{id}$, we can ge the data table information from the
metadata  $\langle A_{id}, DTableId\rangle$. 
Then, we get the relevant data table information and the needed attribute
information.

Let $RD$ be the set of relevant data tables and $d_{a_1} \cdots d_{a_m}$
are the related attributes for table $d\in RD$. We can translate the original OM query to SQL queries
over the data tables.

[
{\tt 
SELECT DISTINCT d.record\_id\\
FROM d\\
WHERE $f(d_{a_1} \cdots d_{a_m}) = true$\\
$[$GROUP BY$]$\\
$[$HAVING$]$
}
]

The above describes the process for the most basic OM query. 
However, in the real application, people ask queries with distinct
observation or entity constraints. Also, people ask queries about
different observatitons using context.

%Discuss how to deal with the distinct constraints.
Many times, a user may want to find {\em distinct} observations or
entities. In this case, the above simple solution cannot work
correctly because it treats observations in different rows as the same
observation. So, we need to use the annotation to figure out what
distinguishes one observation from another. 
When an observation type is denoted with {\em distinct yes} and some
of its measurement types are denoted with {\em key yes}, 
the values of the measurements on the key measurement types can
uniquely distinguish one observation from the other. 
To implement this, in SQL, we can perform ``group by'' operations on
the key measurement types of the observations.

%Discuss how to deal with context
The above description discusses cases without context. 
However, in real application, the observation has context. 
When the context is not {\em identifying}, i.e., the context value
does not affect the uniqueness of one observation, this can be
processed as the basic way. 
When the context is denoted with {\em identifying yes}, it means the
identity of one observation also depend on the contextual
observation. 
In this case, we need to figure out the all the key measurement types
of one observation, then the ``GROUP BY'' operation should be on all
these key measurement types.

{\bf TODO: put an example here.}

In this case, there is no need to change the SQL scripts.
the only change are the relevant attribute set $d_{a_1} \cdots
d_{a_m}$, which corresponds to not only the measurement types that
directly related to the observation, but also the measurement types
that are in the identifying context chain of the observation.

\begin{example}\label{eg:query_reqerite_q1}
{\bf TODO: refine this example} 
Take $Q_1$ as example. The {\tt concept} is ``observation'' and {\tt cond} is {\em species = ``Picea rubens''};

From the annotation $A.MeasType$, we can find the measurement type
$m_3$ with the characteristic {\em species}. 
From the mapping $A.Map$, we further find that the attribute {\em spp} that the
measurement type $m_3$ is mapped to.

The second step is to find the data file which really have this value. 
For this one, we find the data table, and do a selection on the table
content.

{\bf TODO: SQL here.} 
\end{example}

From the above description, we can tell that the computation cost is to search the different data
tables. This way, for each candidate data table, we need to send an
SQL query to the server for evaluation purpose.
If the number of candidate data tables are small. This method should
work well. However, when we have a lot of candidate data tables, the
efficiency may be affected. 
In what follows, we introduce another strategy, which logically merge
the data content in different tables and perform queries ober the
materialized database.

\subsection{Querying over materialized database}\label{sec:matdb}

The previous section shows the method to rewrite a given query. 
However, as we analyzed, the computation cost may be higher when the
candidate data tables are many. In this section, we propose another
strategy to make use of the existing optimization strategies of
current DBMS.

The basic idea is to materialize the data into some centralized
concept tables with instances of entities, observations, measurements
and the contexts between observations. 
We call such concept tables core OBOE tables, and denote them as
$OBOE.*$. 
In partibular, we have the following concept instance tables. 
\begin{itemize}
\item $OBOE.EI$ = $\{(\underline{ei_{id}}, et_{id})\}$ for all the entity instances;
\item $OBOE.OI$ = $\{(\underline{oi_{id}}, ot_{id}, ei_{id})\}$ keeps
  all the observation instances;
\item $OBOE.MI$ = $\{(\underline{mi_{id}}, oi_{id}, mt_{id},
  mVal)\}$ for all the measurement instances;
\item $OBOE.CI =\{(\underline{oi_{id},coi_{id},ct_{id}})\}$ for
  all the context instances;
\end{itemize}

In what follows, we describe how to materialize the data into such
concept instance tables in Section \ref{sec:materialilzedb} when
considering all the distinct and context constratins. 
Then, in Section \ref{sec:materialilzedb}, we describe our strategy to
perform queries over materialized database.

\subsubsection{Materialize data}\label{sec:materialilzedb}
{\bf: TODO: refine the writing in MaterializeDB and move them here.}

\subsubsection{Querying materialized database}\label{sec:materialilzedb}

\begin{example} \label{eg:materialize_db_q1}
Example \ref{eg:query_reqerite_q1} shows how to deal with $Q_1$ using
the query rewriting method. 
Using the  materialized database, the first step is the same.
After we know the data file, then we search the OBOE.MI with the
condition that {\em mVal=``Picea rubens''}.
\end{example}

{\tt 
SELECT distinct $AD.d_{id}$\\
FROM MI, MT, AD\\ 
WHERE $MT.Cha='Species'$ \\
AND $MI.mVal = `Picea rubens'$\\
AND $MI.mt_{id}=MT.mt_{id}$ AND $MT.A_{id}=AD.A_{id}$; %\#Join the MI and MT tables
}

In this example, the computation cost is the selection cost of the
measurement instance table. 
It should be faster than the query rewriting.

If the query has some restriction on the aggregate result, we can
easily add the GROUP BY and HAVING clause to deal with it. 
We take $Q_3$ as an example, which asks for data sets that contain at
least five distinct ``Picea rubens'' observations.
The SQL query can be written as: \\
{\tt 
SELECT distinct $d_{id}$\\
FROM MI, MT\\ 
WHERE $MT.Cha='Species'$ \\
AND $MI.mVal = `Picea rubens'$\\
AND $MI.mt_{id}=MT.mt_{id}$\\
GROUP BY $d_{id},oi_{id}$\\
HAVING COUNT(*)$>$5; %\#Join the MI and MT tables
}

Using the query rewriting, 
First, we find the data tables that contain the needed concepts. 
In each table, we need to figure out the distinct observations. 
From $m_i$ $obstype_i$, all the other measurements that are of
$obstype_i$. 
Group by the key measurements of $obstype_i$ if distinct yes.
If this observation type does not have any context, then the key
measurements are all the measurements directly defined under this
observatioin type. Otherwise, i.e., this observation type has context,
then the key measurements are all the measurements in the context
chain. 
%\end{example}

When we are given contextualized query, e.g., $Q_2$. 
Give me the datasets that contain species ``Picea rubens'' observations in ``California''.

Species = ``Picea rubens'' and observation has context {\em in}
another observation which has measurement of value ``California''.

%Using materializedDB, we can answer the query this way
%First, from the Characteristic ``Species'', we can get  --> mlabel, observation type

%select mtype, olabel from measurement\_type  \\
%where chracteristic = 'Species';  \\

{\tt 
SELECT record\_id\\
FROM CI as ci, MI as mi, MI as cmi,MT as mt, MT as cmt, \\
WHERE $mi.oi_{id}=ci.oi_{id}$ AND $cmi.oi_{id}=ci.coi_{id}$ \\
AND $ci.coi_{id} = coi.oi_{id}$ AND oi.otype = 'olabel' \\
AND mi.mvalue=`Picea rubens' AND cmi.mvalue='California'\\
AND MT.mt_{id}=mi.mt_{id} AND MT.mLabel = 'Species'\\
AND CMT.mt_{id}=cmi.mt_{id} AND cmt.mLabel=`State'\\
AND coi.otype IN (\\
\verb|    |SELECT context\_olabel\\
\verb|    |FROM context\_relationship\\
\verb|    |WHERE olabel = 'olabel');
}
The sub-query finds the set containing all the observation type
labels in the context chain.

Use query rewriting, what I can do? 
It's the same till to getting context\_olabel\_set.

Get the asked condition columns and the context columns,

select record\_id from data\_tables
where condition column 1='value1' and (context column1 = 'context
value' or context column2 = `context value', etc.);

\begin{example}
SQ2: Give me the datasets that have measurements with average ``area''
bigger than 5.0 square feet. 
\end{example}

Use materialized database: 
Analysis: ``Area'' is the context of some observation.
Area --> mlabel, observation type

From the context observation lable, find the observation labels, \\
select olabel\\
from context\_relationship\\
where context\_olabel = 'olabel';

Let olabel\_set  be the set containing all the observation type
labels that use ``Area'' as context.

SELECT oi.record\_id \\
FROM observation\_instance oi, measurement\_instance mi, 
observation\_instance coi, measurement\_instance cmi,context\_instance ci\\
WHERE (mi.oid=ci.oid AND ci.oid = oi.oid) and (cmi.oid=coi.oid AND ci.context\_oid =
coi.oid) AND coi.otype $\in$ olabel\_set 
AND cmi.mlabel=`Area'
GROUP BY cmi.oid
HAVING avg(cmi.mvalue)>5.0in;

\section{Experiments}

Synthetic data generator.

Algorithm to materialize DB. 
Report the time and space.

Synthetic query generator.

Test algorithm to perform query over materiazed DB.
Test algorithm to perform query-rewriting over materiazed DB.

Test algorithm for half-materialized data.

\section{Conclusion}

\bibliographystyle{IEEEtranS}
\bibliography{IEEEabrv,DataIntegration}

\end{document}