### 2011-05-26 docs/pubs/EIM-2011/summary.tex

\section{Summary}
\label{sec-summary}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: