### 2011-05-26 docs/pubs/EIM-2011/annotation.tex
add more sections (for me)

\section{Annotation}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: