### 2011-06-02 docs/pubs/EIM-2011/intro.tex

\section{Introduction}
\label{sec-intro}

A major challenge in environmental information management concerns
providing effective approaches for the discovery and integration of
heteregenous data sets. For instance, locating and combining relevant
observational data are often critical and time-consuming steps for
researchers studying phenomena at broad spatial, temporal, and
biological scales
\cite{worm06:_impac_of_biodiv_loss_ocean_ecosy_servic,pennings05:_do,green05:_compl_in_ecolog_and_conser,sorokina09:_detec_inter_variab_inter_obser_ornit_data,jackson01:_histor_overf_and_recen_collap}. The
underlying data sets used within such studies frequently differ in
subtle and complex ways, due in part to the protocols used for data
collection, the types of observations made, and the experimental and
other contextual information associated with the data set. These
differences in turn can lead to structural and semantic heterogeneity
among data sets that make them hard to discover using current data
management approaches and require considerable manual effort by
researchers needing to combine data sets.

A number of recent efforts within the earth and environmental
informatics communities are adopting the notion of an observation as a
key modeling concept for enabling improved discovery and integration
of scientific data
\cite{om,fox09:_ontol,tarboton07:_cuahs_commun_obser_data_model,cushing07:_compon_based_end_user_datab,balhoff10:_phenex,bowers08}. These
approaches provide higher-level observational data models for
describing and representing observations and measurements found in
underlying datasets by defining common ``core'' concepts such as the
entities or features being observed, measurement units and protocols,
and context relationships between observations \cite{om,bowers08}.  A
major goal of these approaches is to enable interoperability and
uniform access to data by abstracting away the underlying
representation details that often exists across scientific data sets.

In this paper we describe extensions to the Ecological Metadata
Language (EML) \cite{} and suppporting tools for enabling improved
discovery and integration of ecological data sets. Our work is based
on the Extensible Observations Ontology (OBOE) \cite{}, which
represents a generic observational model implemented in OWL-DL for
describing domain-specific observation and measurement types. Our
approach adds additional metadata in the form of semantic annotations
that link attributes within datasets to OBOE terms for describing the
implicit observation and measurement types found within data
sets. Semantic annotations are executable in the sense that they can
be used to convert a data set into a collection of observation and
measurement instances, providing a more uniform representation for
expressing queries and performing integration.  To support the
creation of annotations, we have extended the Morpho metadata editor
with a high-level user interface as well as the Metacat data catalog
for storing and querying annotations through a new Semantic Mediation
API. This API can also be used to perform basic data-level integration
tasks using our prior work on the EML Data Manager Library \cite{}.

The rest of this paper is organized as follows. \secref{sec-framework}
briefly describes the various components used within our
approach. \secref{sec-annotation} describes the extensions we have
developed for Morpho and Metacat to support semantic
annotation. \secref{sec-applications} describes the types of data
discovery queries and integration services supported by our
framework. \secref{sec-related} briefly describes related work, and we
summarize our contributions in \secref{sec-summary}.

%\section{Preliminaries}

%Briefly describe EML, Metacat, and Morhpo. Briefly describe oboe and
%semantic annotations.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: