### 2011-06-02 docs/pubs/EIM-2011/summary.tex

\section{Summary}
\label{sec-summary}

% By providing mechanisms for linking datasets to ontological terms
% organized by a high-level observational model, these new extensions to
% Metacat and Morpho help to overcome a number of limitations in
% existing metdata management systems for data discovery and
% integration.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: