### 2011-06-02 docs/pubs/EIM-2011/integration.tex
integration draft

\section{Integration}

The materialization routine used for semantic data queries laid the groundwork for our ongoing data integration pursuits.
In addition to inspecting the data for values within a range and returning the packages that contain a match, the "smart" data integration feature goes further by constructing a synthetic data product that represents the complete results of the query both in terms of the attributes and the values that are returned. Each original data package match may have wildly different syntactic structures (e.g. column number, naming, order) but could share a subset of attributes that are asserted as semantically compatible. These compatible, in-common attributes then become the facets and metrics in the synthetic dataset with their values filtered accordingly. While preliminary results have been quite promising - albeit academic - continued development of algorithms for determining compatibility of annotated measurements (e.g. gram and ounce are both mass units) and for mutating measurement values using ontologically-defined conversions (e.g. 1000 milligrams in a gram) will enable more thorough data integration.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: