### 2011-06-03 docs/pubs/EIM-2011/annotation.tex
OBOE/Annotation figure for Kelp CHN dataset. needs better caption

\section{Annotation}
\label{sec:annotation}

Semantic Annotations describe the observation model contained in a data object. The Annotation serves as a template for constructing a fully fleshed OBOE model of the [usually tabular] data. Less formal descriptions of the data attributes are often included in traditional metadata formats like EML. while there is some overlap between these two mechanisms - particularly with respect to measurement standards or units - semantic concepts applied to each attribute have the benefit of placing the descriptive burden on the ontology from which the concepts are drawn. Because the Annotation represents a potentially subjective perspective on the data, they are stored independently as XML; the former referencing the latter.

The Annotation structure largely mirrors the core classes in OBOE. Observations are composed of Measurements of a specific Entity and are represented by a collection of Characteristics (usually just one) collected using a defined Protocol and Standard (i.e. unit). 
Tablular data object attributes are mapped to a Measurement such that a collection of attributes usually pertain to the Entity of a shared Observation (\figref{fig:kelp-mass-model}). These Observtions can provide context for other Observations such that the structure of the observational data model and collection paradigm are formally represented in the Annotation.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{images/kelp-mass-model.png}
\caption{Partial OBOE Annotation for Kelp sampling data. Shaded nodes represent ontological concepts; rectangular nodes are data table attibutes mapped to OBOE characteristics.}
\label{fig:kelp-mass-model}
\end{figure}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: