### 2011-06-04 docs/pubs/EIM-2011/summary.tex

\section{Summary}
\label{sec:summary}

The Semtools project has been successful in exploring and codifying technologies and techniques for applying semantic concepts to observational data. By providing mechanisms for linking datasets to ontological terms organized in a high-level observational model (e.g., OBOE), these new extensions to Metacat and Morpho help to overcome a number of limitations in
existing metadata management systems that strive to provide effective data discovery and integration features. Our close involvement with the SONet Project (Scientific Observations Network) \cite{sonet} encourages continued use-case refinement that will inform future semantic tool development and place an emphasis on intuitive interfaces and incremental adoption. This varied community of stakeholders is firmly invested in the use of cutting edge semantic solutions that will ultimately benefit multiple science disciplines by reducing obstacles to broad data sharing and innovative reuse.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: