### 2011-06-05 docs/pubs/EIM-2011/summary.tex

\section{Summary}
\label{sec:summary}

The Semtools project has been successful in exploring and codifying
technologies and techniques for applying semantic concepts to
observational data. By providing mechanisms for linking data sets to
ontological terms organized in a high-level observational model (e.g.,
OBOE), these new extensions to Metacat and Morpho help to overcome a
number of limitations in existing metadata management systems that
strive to provide effective data discovery and integration
features. Our close involvement with the SONet Project (Scientific
Observations Network) \cite{sonet} encourages continued use-case
refinement that will inform future semantic tool development and place
an emphasis on intuitive interfaces and incremental adoption. This
varied community of stakeholders is firmly invested in the use of
cutting edge semantic solutions that will ultimately benefit multiple
science disciplines by reducing obstacles to broad data sharing and
innovative reuse.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: