### 2011-07-30 docs/pubs/EIM-2011/framework.tex
snapshot copy -- will finish up rest later

\section{Semtools Framework}
\label{sec:framework}

The Semtools project has focused development efforts on three main
components: a Java library for accessing and manipulating OBOE
ontology extensions and semantic annotations, an annotation plugin for
the Morpho metadata editor, and query extensions for the Metacat data
catalog. Below we describe these components and give a brief
overview of the OBOE model as well as the semantic annotation approach
used in Semtools. For a more in-depth presentation of OBOE see
\cite{madin07:_ontol_for_descr_and_synth,bowers08}.

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.494\textwidth]{images/oboe}
  \caption{Main classes and properties of the extensible observation
    ontology (OBOE). While shown using the Unified Modeling Language (UML),
    the model is defined as an OWL-DL ontology.}
  \label{fig:oboe}
  \vspace{-12pt}
\end{figure}

\subsection{The OBOE Observational Model}

\figref{fig:oboe} shows the main modeling constructs of OBOE (see:
\url{http://ecoinformatics.org/oboe/oboe.1.0/oboe-core.owl}). An {\em
  observation} is made of an {\em entity} (e.g., biological organisms,
geographic locations, environmental features) and serves to group a
set of measurements together to form a single ``\emph{observation
  event}''. A \emph{measurement} assigns a value to a {\em
  characteristic} of the observed entity (e.g., the weight of a
plant), and can also include \emph{standards} (e.g., units as well as
standards for coded values) and collection \emph{protocols}. An
observation can occur within the surrounding \emph{context} of other
observations (e.g., as part of a temporal or spatial context), and
context may include a named relationship (e.g., ``partOf'',
``within'') that existed during the observation event. A key feature
of OBOE is that it allows properties (characteristics and
relationships) of entities to be asserted without being interpreted as
\emph{inherently} (i.e., {always}) true of the entity.  Depending on
the context in which the entity was observed or how the measurements
were performed, an entity's properties may take on different values.
OBOE allows RDF-style assertions about entities to be contextualized,
and thus different values can be assigned for the same entity under
distinct contexts, which is a crucial feature for modeling ecological
as well as many other types of scientific data
\cite{bowers08,mungall07:_repres_phenot_in_owl}. In addition, OBOE is
currently implemented as an OWL-DL ontology that can be easily used
with (or extended by) other ontologies for specifying domain-specific
types of entities, characteristics, measurement standards, protocols,
and relationships. For instance, the Semtools project has defined
specific OBOE extensions in collaboration with the Santa Barbara
Coastal Long-Term Ecological Research Project as well as through
ongoing collaborations with other projects, and general extensions
exist for OBOE that define a number of common entities, measurement
units, and corresponding physical characteristics.

\begin{figure}[!b]
\centering
\includegraphics[width=0.5\textwidth]{images/kelp-mass-model.pdf}
\caption{Partial OBOE semantic annotation for Kelp sampling
  data. Shaded nodes represent ontological concepts; rectangular nodes
  are data table attributes mapped to OBOE measurement
  characteristics.}
\label{fig:kelp-mass-model}
\end{figure}

\subsection{Semantic Annotations}

A semantic annotation consists of two parts: (i) a ``configuration''
of the observation model containing the specific entities,
characteristics, observations, measurements, and so on (drawn from one
or more domain ontologies) that appropriately capture the semantics of
the data set; and (ii) a mapping between the attributes in the data
set to specific measurements defined in the model
configuration. \figref{fig:kelp-mass-model} shows a high-level example
of an annotation defined for a simple Kelp sampling data set. Here,
the data set consists of five attributes (bottom of
\figref{fig:kelp-mass-model}). Each attribute is mapped to a specific
measurement type (where only the characteristic of each measurement
type is shown), and measurement types are organized into observations
of specific Kelp entities (shown of type ``Macrocystis''), temporal
points (denoted by date-times), and spatial locations (given as site
names). Each measurement associated with a Kelp observation is assumed
to have occurred within the site and during the given time as
specified by the context relationships.

\begin{figure*}[!t]
\centering
\includegraphics[width=1.0\textwidth]{images/morpho-annotation-widget.png}
\caption{Morpho metadata editor with Semantic
  plugin. The fill-in-the-blank interface uses
  natural language descriptions for intuitive editing. A searchable, hierarchical browser is
  used to select concepts from domain-specific ontologies.}
\label{fig:morpho-annotation}
\end{figure*}

Semantic annotations can be used to facilitate discovery and
integration of heterogeneous data sets. For instance, combining
semantic annotations with OBOE, it is possible to discover data sets
based on searches expressed over types of observations and
measurements of interest. As simple examples, users can pose queries
such as ``\emph{find all data sets containing observations of Kelp}''
and ``\emph{find all data sets containing Mass measurements of
  Kelp}''.  Both of these queries would return the example data set in
\figref{fig:kelp-mass-model} since the attribute {\tt WET} is linked
to a WetMass measurement for observations of Macrocystis (where
WetMass is defined as a special kind, or \emph{subclass} of Mass, and
Macrocystis is defined as a subclass of Kelp). Using semantic
annotations in this way can help to increase both query recall and
precision over standard keyword-based approaches
\cite{berkley09:_improv_data_discov_for_metad}. In particular, by
defining terms as subclasses of other terms (e.g., Macrocystis as a
subclass of Kelp), term expansion can be used to increase the number
(recall) of data sets returned (where subclasses of query terms are
also searched). The precision of the result can be improved since
queries may specify the desired connections between terms (e.g.,
measurements of Mass for Kelp observations) as opposed to returning
all data sets that simply mention the terms but without any explicit
connections (i.e., where Mass was measured, but not for Kelp samples).
% Queries can additionally include context constraints, e.g., Mass
% measurements of Kelp within a specific spatial or temporal context.
Annotations also help facilitate integration by allowing
tools to align data set attributes based on their declared measurement
and observation types.

In general, semantic annotations provide a formal description of
attribute semantics, whereas in many commonly-used metadata formats
for describing data sets, only informal text-based descriptions of
data attributes are permitted.  In the case of EML, there is some
overlap between these two mechanisms---particularly with respect to
measurement standards---however, semantic annotations extend this
approach by providing a general mechanism to formally associate
concepts drawn from domain ontologies to attributes. In the Semtools
framework, we employ an XML serialization syntax for semantic
annotations that is compatible with EML but that is stored separately
from the EML documentation of a data set (allowing, e.g., annotations
to be used independently of EML or with other metadata standards if
needed). In addition, semantic annotations can be used to
``materialize'' a given data set into a set of triples conforming to
the model configuration given in the annotation. In other words, a
tabular data set such as the one shown in \figref{fig:kelp-mass-model}
can automatically be converted into a corresponding collection of
observation and measurement instances. This in turn enables a simple
form of structural integration, where instead of having a large number
of different tabular data structures, all data is represented using
the standard set of structures defined by the OBOE model (see
\figref{fig:oboe}). Thus, materializing a data set in this way
provides a more uniform structural representation that can make a
number of discovery and integration tasks easier. For instance,
materialization can be used to increase query expressivity by allowing
searches of the form ``\emph{find all data sets containing Mass
  measurements of Kelp with values less than or equal to 5 grams}'', which in
our example can be answered by generating (i.e., materializing) the
measurements associated with the {\tt WET} and {\tt DRY} attributes in the data set
of \figref{fig:kelp-mass-model}.

% Because the Annotation represents a potentially subjective
% perspective on the data, they are stored independently as XML; the
% former referencing the latter.  The Annotation structure largely
% mirrors the core classes in OBOE. Observations are composed of
% Measurements of a specific Entity and are represented by a
% collection of Characteristics (usually just one) collected using a
% defined Protocol and Standard (i.e. unit).  Tablular data object
% attributes are mapped to a Measurement such that a collection of
% attributes usually pertain to the Entity of a shared Observation
% (\figref{fig:kelp-mass-model}). These Observtions can provide
% context for other Observations such that the structure of the
% observational data model and collection paradigm are formally
% represented in the Annotation.

\subsection{The Semantic Mediation API}

The Semantic Mediation API includes basic ontology management
features, annotation manipulation capabilities, and simple concept
navigation and visualization components. The API is intended to be a
centralized toolkit for use in multiple application contexts (on
either client or server deployments).  The Semantic Mediation API uses
both the OWL API \cite{owlapi} for ontology management services
including ontology parsing, serialization, and simple class and
property exploration as well as the Pellet description-logic reasoner
\cite{pellet} for classification and exposing inferred axioms in
source ontologies. The inference services exposed through the Semantic
Mediation API are used in both discovery and integration features
described below.
% is particularly powerful when utilizing OBOE Measurement 'templates'
% that express exactly what (Entity and Characteristic) can be
% observed and how (Standard and Protocol).
In our current Morpho and Metacat extensions, semantic annotations are
managed and stored automatically in an underlying, local relational
database. While it is also possible to use in-memory approaches for
storing and querying annotations, we found the overhead to be
prohibitive when large numbers of data sets are managed.

% store and query annotations inmemory
% approachesfor storing for scalable persistence and improved query
% performance, which we found to be essential when large numbers of data
% sets are supported. , where the overhead of in-memory annotation
% queries having proven to be prohibitively limiting for large sets of
% data.

\subsection{The Morpho Editor Plugin}

The semantic-annotation editor plugin for Morpho provides a front-end
to the Semantic Mediation API and allows data owners and curators to
define annotations for existing EML data descriptions. The editor
provides a simple ``fill-in-the-blank'' style form-based interface
with a searchable hierarchical concept selection widget (see
\figref{fig:morpho-annotation}). The plugin seamlessly integrates with
a standard Morpho installation and provides semantic query
capabilities for locating data packages, marking up data sets within a
package using semantic annotations, and saving annotations locally or
to a shared repository where they can be discovered and explored by
other users. The annotation editor in Morpho allows a user to view the
data set being annotated as they fill in (by selecting an appropriate
ontology term) the characteristic, measurement standard, protocol, and
associated entity for each data set attribute. Users can also specify
whether an observation spans multiple columns, and can provide context
relationships between attributes (i.e., observations). The editor
provides a number of additional features including the ability to view
the entire annotation (similar to \figref{fig:kelp-mass-model}) and to
specify additional mapping constraints for observations and
measurements.

\subsection{Metacat Query Extensions}

The semantic plugin for Metacat augments Metacat's existing metadata
storage and search by allowing annotations to be saved and queried
alongside the metadata and data that they annotate. In addition to
traditional keyword and spatial search criteria, the Metacat plugin
allows semantic criteria to be included where they may either increase
query recall using term-expansion (i.e., traversing the class
subsumption hierarchy) or refine the result set by limiting matches to
data sets that contain the specified observational model (e.g.,
combinations of OBOE-compatible entity, characteristic, measurement
standard, or protocol concepts). The observational model can be
leveraged further by materializing the annotation and data artifact
(via the Data Manager Library
\cite{leinfelder10:_metad_driven_approac_to_loadin}) into a fully
instantiated OBOE model and inspecting (and querying over) the
observational values themselves.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: